Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-03-19 ltx/BrainIDE.tex
further updated paper following meeting on the 20/03/2010

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behaviour. 
\begin{todo}
	{Really need to add some citations here.}
\end{todo}
It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. 
\begin{todo}
	{Same here.}
\end{todo}
The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology. The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at position $r$ to the local post synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations at locations $\mathbf{r}'$ can be described by
\begin{equation}\label{SpikesToPotential}
	v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}.
\end{equation}
The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
	h(t) = \eta(t)\exp{\left(-\alpha t\right)}.
\end{equation}
where $\alpha=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. The synaptic response kernel can be more elaborate, using rise and fall time constants, however, this simple form serves our purposes. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
	g\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'} + u\left(\mathbf{r},t\right)
\end{equation}

@@

	v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'

@@

	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t

@@

\section{Derivation of Finite Dimensional State-Space Model}
In order to implement standard estimation techniques we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
	v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t
\end{equation}
where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. The connectivity kernel can also be considered a decomposition into a set of basis functions in 
\begin{equation}\label{DefKernelDecomp}
	w\left(\mathbf{r}-\mathbf{r}'\right) = \boldsymbol{\theta}^{\top} \mathbf{\psi}\left(\mathbf{r}-\mathbf{r}'\right)
\end{equation}
We will assume we know the parametric form of the connectivity basis functions, where the parameter $\boldsymbol{\theta}$ is unknown. Each connectivity basis function can be considered a layer in the Wilson and Cowan model, representing short range excitation, surround inhibition and mid-range excitation. Making substitutions of~\ref{DefFieldDecomp} and~\ref{DefKernelDecomp} into~\ref{DiscreteTimeModel} we get
\begin{equation}\label{reduced continuous model}
	\boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T\boldsymbol{\theta}^{\top}\int_\Omega  {\psi \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'}  - \lambda\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right).
\end{equation}
By defining the term
\begin{equation}\label{DefGamma}
	\boldsymbol{\Gamma}  = \int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)d\mathbf{r}}
\end{equation}
and manipulating the model (see Appendix~\ref{Simplifying Decomposition}) we isolate the state vector to form a state-space model
\begin{eqnarray}\label{Homogeneous SS Model}
	\mathbf{x}_t &=& q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) &=& T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\psi^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \lambda \mathbf{x}_t \\
	k(u_t(\mathbf{r})) &=& \boldsymbol{\Gamma}^{-1}T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} \\
	\mathbf{e}_t &=& \boldsymbol{\Gamma}^{-1}\int_\Omega{\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}} \\
\end{eqnarray}
where $\mathbf{e}_t$ is a zero-mean normal white noise term with the covariance (see Appendix D)
\begin{equation} 
\boldsymbol{\Sigma}_w=\boldsymbol{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \rho\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 
\end{equation}

\begin{equation}
 cov(\mathbf{e}_t,\mathbf{e}_{t+\tau})=
\begin{cases}
\rho\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
0 & \mathrm{otherwise}
\end{cases}
\label{eq:FieldCovariance}
\end{equation}

The observation equation of the reduced model is
\begin{eqnarray}\label{ObservationEquation}
    \mathbf{y}_t &=& z\left(\mathbf{x}_t\right) + \boldsymbol{\varepsilon}_t \\
	z\left(\mathbf{x}_t\right) &=& \int_\Omega m\left(\mathbf{r}_n-\mathbf{r}'\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) d\mathbf{r}' \mathbf{x}_t.
\end{eqnarray}

@@

\section{Discussion}

@@

\section{Discrete Time Model}\label{Time Discretization}
To form the IDE neural field model a time discretization must be perform. We used a one-step Euler method where Eq.~\ref{FinalForm1} can be approximated by
\begin{equation}\label{Euler Approximation}	
\frac{v\left( \mathbf{r},t+T \right) - v\left( \mathbf{r},t\right)}{T} =   -\alpha v\left( \mathbf{r},t \right) + \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'} + u\left(\mathbf{r},t\right).
\end{equation}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging Eq.~\ref{Euler Approximation} we get
\begin{eqnarray}\label{Euler Approximation}	
v_{t+1}\left( \mathbf{r}\right) &=& v_t\left( \mathbf{r}\right) -T \alpha v_t\left( \mathbf{r}\right) + T \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'} + T u_t\left(\mathbf{r}\right)
\end{eqnarray}
The discrete time form of the model is
\begin{equation}\label{Discrete Time Model1}
	v_{t+1}\left(\mathbf{r}\right) = \lambda v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right).
\end{equation}
where $\lambda = 1 - T \alpha$.

\section{Numerical Solution to IDE Model}\label{Space Discretization}
To solve the intregro-difference equation in Eq.~\ref{Discrete Time Model1} we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta $ giving
\begin{equation}\label{discrete space}
	v_{t+1}\left(\mathbf{r}_{ij}\right) = \lambda v_t\left(\mathbf{r}_{ij}\right) + T \Delta^2 \sum_{i=1}^{i=I}{ \sum_{j=1}^{j=J}{ w\left( \mathbf{r}-\mathbf{r}_{ij}' \right)f\left( v_t\left( \mathbf{r}_{ij}'\right) \right)} } + T u_t\left(\mathbf{r}\right) + e_t(\mathbf{r}_{i,j})
\end{equation}
where $e_t(\mathbf{r}_{i,j}) \sim \mathcal{N}\left(0,\Sigma\right)$.
%  and $\tilde{w}(r_{ij}-r_{ij}'|)$ discretized connectivity kernel described by
% \begin{equation}\label{Discrete Connectivity Kernel}
% 	\tilde{w}(r_{ij},r_{ij}') = \boldsymbol{\theta}^{\top} e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
% \end{equation}
% The observation equation and kernel were discretized in the same way by
% \begin{equation}\label{Discrete Observation Equation}
%     y_k\left(r_{i_n j_n}\right) = \Delta^2\sum_{i=1}^{i=I}\sum_{j=1}^{j=J} \tilde{m}( |r_{i_n j_n}-r_{ij}'| ) v_k\left(r_{ij}'\right)  + \varepsilon_k( r_{i_n j_n})
% \end{equation}
% where
% \begin{equation}\label{Discrete Observation Kernel}
% 	\tilde{m}(|r_{ij}-r_{ij}'|) = e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
% \end{equation}
We have used periodic boundary conditions to alleviate problems associated with edge effects.

\section{Reduction to Finite State-Space Model}\label{Simplifying Decomposition}
Starting from Eq.~\ref{reduced continuous model} we multiply throughout by $\boldsymbol{\phi}(r)$ and integrate over the spatial domain $\Omega$ to get
\begin{eqnarray}\label{Poo}
	\int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) d\mathbf{r}} \mathbf{x}_{t+1} = T \int_\Omega  {\boldsymbol{\phi} (\mathbf{r}) \boldsymbol{\theta}^{\top} \int_\Omega  {\psi \left(\mathbf{r}-\mathbf{r}'\right)f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right) \mathbf{x}_t \right)d\mathbf{r}'}d\mathbf{r}} - \lambda\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\phi}^{\top}(\mathbf{r})d\mathbf{r}} \mathbf{x}_t \\
	+ T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} + \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) e_t\left(\mathbf{r}\right)d\mathbf{r}}.
\end{eqnarray}
Substituting Eq.~\ref{DefGamma} into Eq.~\ref{Poo} and cross-multiplying by $\boldsymbol{\Gamma}^{-1}$ gives 
\begin{eqnarray}\label{Homogeneous SS Model}
	\mathbf{x}_{t+1} = T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\psi^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \lambda \mathbf{x}_t \\
	+ \boldsymbol{\Gamma}^{-1}T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} + \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}}.
\end{eqnarray}

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{Colored noise}
We have
\begin{equation}
v_{t + 1}\left( \mathbf{r} \right)  = T\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}} + \lambda v_t\left( {\bf{r}} \right) +e_t\left( {\mathbf{r}} \right).
\end{equation}
where the disturbance $e_t(\mathbf r)$ is a zero-mean normally distributed noise process, spatially coloured but temporally white, with covariance

>>>>

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behaviour. 
\begin{todo}
	{Really need to add some citations here.}
\end{todo}
It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. 
\begin{todo}
	{Same here.}
\end{todo}
The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology. The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at position $r$ to the local post synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations at locations $\mathbf{r}'$ can be described by
\begin{equation}\label{SpikesToPotential}
	v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}.
The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
	h(t) = \eta(t)\exp{\left(-\alpha t\right)}.
where $\alpha=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. The synaptic response kernel can be more elaborate, using rise and fall time constants, however, this simple form serves our purposes. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
	g\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'} + u\left(\mathbf{r},t\right),

@@

	v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'.

@@

	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t,

@@

\section{Derivation of Finite Dimensional State-Space Model}
In order to implement standard estimation techniques we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
	v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t,
\end{equation}
where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. The connectivity kernel can also be considered a decomposition into a set of basis functions in 
\begin{equation}\label{DefKernelDecomp}
	w\left(\mathbf{r}-\mathbf{r}'\right) = \boldsymbol{\theta}^{\top} \boldsymbol{\psi}\left(\mathbf{r}-\mathbf{r}'\right).
\end{equation}
We will assume we know the parametric form of the connectivity basis functions, where the parameter $\boldsymbol{\theta}$ is unknown. Each connectivity basis function can be considered a layer in the Wilson and Cowan model, representing short range excitation, surround inhibition and mid-range excitation. Making substitutions of~\ref{DefFieldDecomp} and~\ref{DefKernelDecomp} into~\ref{DiscreteTimeModel} we get
\begin{equation}\label{reduced continuous model}
	\boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T\boldsymbol{\theta}^{\top}\int_\Omega  {\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'}  - \lambda\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right).
\end{equation}
By defining the term
\begin{equation}\label{DefGamma}
	\boldsymbol{\Gamma}  = \int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)d\mathbf{r}}
\end{equation}
and manipulating the model (see Appendix~\ref{Simplifying Decomposition}) we isolate the state vector to form a state-space model
\begin{eqnarray}\label{Homogeneous SS Model}
	\mathbf{x}_t &=& q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) &=& T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\boldsymbol{\psi}^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \lambda \mathbf{x}_t \\
	k(u_t(\mathbf{r})) &=& \boldsymbol{\Gamma}^{-1}T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} \\
	\mathbf{e}_t &=& \boldsymbol{\Gamma}^{-1}\int_\Omega{\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}},
\end{eqnarray}
where $\mathbf{e}_t$ is a zero-mean normal white noise term with the covariance (see Appendix D)
\begin{equation} 
\boldsymbol{\Sigma}_w=\boldsymbol{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \rho\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 
\end{equation}
\begin{equation}
 cov(\mathbf{e}_t,\mathbf{e}_{t+\tau})=
\begin{cases}
\rho\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
0 & \mathrm{otherwise}.
\end{cases}
\label{eq:FieldCovariance}
\end{equation}
The observation equation of the reduced model is
\begin{eqnarray}\label{ObservationEquation}
    \mathbf{y}_t &=& z\left(\mathbf{x}_t\right) + \boldsymbol{\varepsilon}_t \\
	z\left(\mathbf{x}_t\right) &=& \int_\Omega m\left(\mathbf{r}_n-\mathbf{r}'\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) d\mathbf{r}' \mathbf{x}_t.
\end{eqnarray}

@@

\section{Discussion}
\begin{itemize}

	\item Discuss the novelty of using the reduced model (BrainIDE) for creating patient specific models.

	\item Spatial frequency of field and the relationship to the required density of electrodes/sensors required for estimation or to capture the required information without aliasing.

	\item The flexibility of the approach

	\item Combining this method with other to incorporate longer range hetrogeneous connectivity.

	\item The need for activity to perform estimation.
\end{itemize}

@@

\section{Discrete Time Model}\label{Time Discretization}
To form the IDE neural field model a time discretization must be perform. We used a one-step Euler method where Eq.~\ref{FinalForm1} can be approximated by
\begin{equation}\label{Euler Approximation}	
\frac{v\left( \mathbf{r},t+T \right) - v\left( \mathbf{r},t\right)}{T} =   -\alpha v\left( \mathbf{r},t \right) + \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'} + u\left(\mathbf{r},t\right).
\end{equation}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging Eq.~\ref{Euler Approximation} we get
\begin{eqnarray}\label{Euler Approximation}	
v_{t+1}\left( \mathbf{r}\right) &=& v_t\left( \mathbf{r}\right) -T \alpha v_t\left( \mathbf{r}\right) + T \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'} + T u_t\left(\mathbf{r}\right).
\end{eqnarray}
The discrete time form of the model is
\begin{equation}\label{Discrete Time Model1}
	v_{t+1}\left(\mathbf{r}\right) = \lambda v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right),
\end{equation}
where $\lambda = 1 - T \alpha$.

\section{Numerical Simulation of IDE Model}\label{Space Discretization}
To solve the intregro-difference equation in Eq.~\ref{Discrete Time Model1} we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta $ giving
\begin{equation}\label{discrete space}
	v_{t+1}\left(\mathbf{r}_{ij}\right) = \lambda v_t\left(\mathbf{r}_{ij}\right) + T \Delta^2 \sum_{i=1}^{i=I}{ \sum_{j=1}^{j=J}{ w\left( \mathbf{r}-\mathbf{r}_{ij}' \right)f\left( v_t\left( \mathbf{r}_{ij}'\right) \right)} } + T u_t\left(\mathbf{r}\right) + e_t(\mathbf{r}_{i,j}),
\end{equation}
where $e_t(\mathbf{r}_{i,j}) \sim \mathcal{N}\left(0,\Sigma\right)$.
%  and $\tilde{w}(r_{ij}-r_{ij}'|)$ discretized connectivity kernel described by
% \begin{equation}\label{Discrete Connectivity Kernel}
% 	\tilde{w}(r_{ij},r_{ij}') = \boldsymbol{\theta}^{\top} e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
% \end{equation}
% The observation equation and kernel were discretized in the same way by
% \begin{equation}\label{Discrete Observation Equation}
%     y_k\left(r_{i_n j_n}\right) = \Delta^2\sum_{i=1}^{i=I}\sum_{j=1}^{j=J} \tilde{m}( |r_{i_n j_n}-r_{ij}'| ) v_k\left(r_{ij}'\right)  + \varepsilon_k( r_{i_n j_n})
% \end{equation}
% where
% \begin{equation}\label{Discrete Observation Kernel}
% 	\tilde{m}(|r_{ij}-r_{ij}'|) = e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
% \end{equation}
We have used periodic boundary conditions to alleviate problems associated with edge effects.

\section{Reduction to Finite State-Space Model}\label{Simplifying Decomposition}
Starting from Eq.~\ref{reduced continuous model} we multiply throughout by $\boldsymbol{\phi}(r)$ and integrate over the spatial domain $\Omega$ to get
\begin{eqnarray}\label{Poo}
	\int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) d\mathbf{r}} \mathbf{x}_{t+1} = T \int_\Omega  {\boldsymbol{\phi} (\mathbf{r}) \boldsymbol{\theta}^{\top} \int_\Omega  {\boldsymbol{\psi}  \left(\mathbf{r}-\mathbf{r}'\right) f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right) \mathbf{x}_t \right)d\mathbf{r}'}d\mathbf{r}} - \lambda\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\phi}^{\top}(\mathbf{r})d\mathbf{r}} \mathbf{x}_t \\
	+ T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} + \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) e_t\left(\mathbf{r}\right)d\mathbf{r}}.
\end{eqnarray}
Substituting Eq.~\ref{DefGamma} into Eq.~\ref{Poo} and cross-multiplying by $\boldsymbol{\Gamma}^{-1}$ gives 
\begin{eqnarray}\label{Homogeneous SS Model}
	\mathbf{x}_{t+1} = T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right) \boldsymbol{\psi}^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \lambda \mathbf{x}_t \\
	+ \boldsymbol{\Gamma}^{-1}T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} + \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}}.
\end{eqnarray}

\section{Colored noise}
We have
\begin{equation}
v_{t + 1}\left( \mathbf{r} \right)  = T\int_\Omega  {k\left( {\mathbf{r}} , {\mathbf{r}'} \right)f\left( {{v_t}\left( \mathbf{r'} \right)} \right)d{\mathbf{r'}}} + \lambda v_t\left( {\mathbf{r}} \right) +e_t\left( {\mathbf{r}} \right).
\end{equation}
where the disturbance $e_t(\mathbf{r})$ is a zero-mean normally distributed noise process, spatially coloured but temporally white, with covariance
